---
title: "Trend_analysis"
author: "S. Sillen"
date: '2022-06-09'
output: html_document
---

Required Packages
```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggpubr)
library(xgboost)
library(feather)
```


# Workflow
```{r}

# 1) Read in sf file ; filter to summer obs > 2 per year ; determine percetage of trophic state observation per year per lake

# 2) Summarise classes across 2 time frames (1984-2005) (2005-2020)

# 3) Identify no change lakes, lakes trending eutrophic, and lakes trending oligotrophic

# 4) Create plot that shows an example of how lakes fit into each of these categories
```


# 1) Read in preds file ; filter to summer obs > 2 per year ; determine percetage of trophic state observation per year per lake
```{r}
preds <- read_feather("C:/Users/samsi/OneDrive - University of Pittsburgh/WesternMountainsChlPreds/Data/preds.feather")

preds <- preds %>%
  mutate(pred = case_when(pred==0~'oligotrophic',
                                pred==1~'mesotrophic',
                                pred==2~'eutrophic')) %>%
  mutate(month = month(date, label = TRUE)) %>%
  mutate(year = as.integer(year)) %>%
  filter(year != 2020) %>% #Limnosat ends in June 2020 so there are very few (4) sites that have enough 2020 to be included here
  filter(type == 'dp',
         month %in% c('Jun', 'Jul', 'Aug', 'Sep')) %>%
  mutate(year = year(date)) %>%
  group_by(Hylak_id, year) %>%
  filter(n() >= 2) %>%
  ungroup()

# Filter for 35 years of data (1984-2019)
df <- preds %>%
group_by(Hylak_id) %>%
mutate(unique_years = length(unique(year))) %>% 
filter(unique_years >= 35) %>% ungroup() 

preds <- preds %>%
  filter(Hylak_id %in% df$Hylak_id)

# get total number of observations at each site for each year so we can calculate percentages later 
yearly_observations <- preds %>%
  group_by(Hylak_id, year) %>%
  summarise(total_obs = n()) %>%
  ungroup()

# prepare for loop
trophic_states <- c('oligotrophic', 'mesotrophic', 'eutrophic')

# for each trophic state, get the number of observations within each year at each site
for (i in trophic_states) {
    state <- preds %>%
      filter(pred == i) %>%
      group_by(year,Hylak_id) %>%
      summarise(count = n())
    yearly_observations <- left_join(yearly_observations, state, by = c('year', 'Hylak_id'))
}

# calculate percentages based on total number of obs in a year and the number of obs in a year for each trophic state 
yearly_observations <- yearly_observations %>%
  mutate_all(funs(replace_na(.,0))) %>%
  rename(count_oligotrophic = 'count.x',
         count_mesotrophic = 'count.y',
         count_eutrophic = 'count') %>%
  mutate(pct_oligotrophic = (count_oligotrophic / total_obs)*100,
         pct_mesotrophic = (count_mesotrophic / total_obs)*100,
         pct_eutrophic = (count_eutrophic / total_obs)*100) 

```

# 2) Summarise (average) classes across 2 time frames (1984-2005 == 'early') (2005-2020 == 'late')
```{r}
early <- yearly_observations %>%
  filter(year < 2005) %>%
  group_by(Hylak_id) %>%
  summarise(early_mean_oligotrophic = mean(pct_oligotrophic), early_mean_mesotrophic = mean(pct_mesotrophic), early_mean_eutrophic = mean(pct_eutrophic))

late <- yearly_observations %>%
  filter(year >= 2005) %>%
  group_by(Hylak_id) %>%
  summarise(late_mean_oligotrophic = mean(pct_oligotrophic), late_mean_mesotrophic = mean(pct_mesotrophic), late_mean_eutrophic = mean(pct_eutrophic))

trends_all <- left_join(early, late, by = 'Hylak_id') %>%
    mutate(oligotrophic_delta = (late_mean_oligotrophic - early_mean_oligotrophic),
          mesotrophic_delta = (late_mean_mesotrophic - early_mean_mesotrophic),
          eutrophic_delta = (late_mean_eutrophic - early_mean_eutrophic))
```

# 3) Identify no change lakes, lakes trending eutrophic, and lakes trending oligotrophic
```{r}
no_trend <- trends_all %>%
  filter(oligotrophic_delta < 10 & oligotrophic_delta >-10, 
         mesotrophic_delta < 10 & mesotrophic_delta > -10, 
         eutrophic_delta < 10 & eutrophic_delta > -10) %>%
  mutate(trend = 'no trend')

trending_eutrophic <- trends_all %>%
  filter(eutrophic_delta > 10, oligotrophic_delta < -10) %>%
  mutate(trend = 'trending eutrophic')

trending_oligotrophic <- trends_all %>%
  filter(oligotrophic_delta > 10 , eutrophic_delta < -10) %>%
  mutate(trend = 'trending oligotrophic')

#bind them up
trends <- rbind(no_trend, trending_eutrophic, trending_oligotrophic)

#Percent of no change lakes
nrow(no_trend) / nrow(trends_all) * 100

#Percent of trending oligotrophic lakes
nrow(trending_oligotrophic) / nrow(trends_all) * 100

#Percent of trending eutrophic lakes 
nrow(trending_eutrophic) / nrow(trends_all) * 100

write.csv(trends, "C:/Users/samsi/OneDrive - University of Pittsburgh/WesternMountainsChlPreds/data/trends.csv", row.names = FALSE)
```

# 4 Explore the general trophic state of the lakes in no trend lakes (i.e. are no chagne lakes mostly oligotrophic or eutrophic)
```{r}
no_trend_all <- yearly_observations %>% filter(Hylak_id %in% no_trend$Hylak_id) %>% 
group_by(Hylak_id) %>% 
summarise(mean_oligo = mean(pct_oligotrophic), mean_eutro = mean(pct_eutrophic))

#Percentage of No trend lakes that are mostly oligotrophic
nrow(no_trend_all %>% filter(mean_oligo > 50)) / nrow(no_trend) * 100 

#Percentage of No trend lakes that are mostly eutrophic
nrow(no_trend_all %>% filter(mean_eutro > 50)) / nrow(no_trend) * 100 

```
# 4) The following code creates Figure 2: Examples of possible trend categories based on the trends in % occurrence of trophic states.  
```{r}
# no trend  

no_trend_example <- no_trend %>%
fiilter(Hylak_id == 1032829) #Johnson Lake Montana

no_trend_example <- yearly_observations %>%
  filter(Hylak_id %in% no_trend_example$Hylak_id) %>%
  rename(Oligotrophic = "pct_oligotrophic", 
         Mesotrophic = "pct_mesotrophic",
          Eutrophic = "pct_eutrophic") %>%
  pivot_longer(cols = c("Oligotrophic", "Mesotrophic", "Eutrophic"), names_to = "Trophic_State")

no_trend_example$Trophic_State <- factor(no_trend_example$Trophic_State, levels = c('Eutrophic', "Mesotrophic", "Oligotrophic"))

no_trend_example_plot <- ggplot(no_trend_example, aes(x = year, y = value, fill = Trophic_State)) +
  geom_bar(stat = "identity", show.legend = FALSE, color = 'black', width = 1) +
  scale_fill_manual(values = c( '#7dae38',  '#698c86', '#2158bc')) +
  labs(title = "A) No trend (Johnson Lake, MT)") +
  xlab("Year") +
  ylab("% Occurence") +
  theme_bw() +
  theme(legend.title = element_blank(), axis.title = element_text(size = 30), axis.text = element_text(size = 24, color = 'black'), legend.text = element_text(size = 30), title = element_text(size = 30, face = 'bold'))

no_trend_example_plot

## increasing in % oligotrophic

trending_oligotrophic_example <- trending_oligotrophic %>%
  filter(Hylak_id == 112052) #Tinmath Reservoir in CO

trending_oligotrophic_example <- yearly_observations %>%
  filter(Hylak_id %in% trending_oligotrophic_example$Hylak_id) %>%
  rename(Oligotrophic = "pct_oligotrophic", 
         Mesotrophic = "pct_mesotrophic",
          Eutrophic = "pct_eutrophic") %>%
  pivot_longer(cols = c("Oligotrophic", "Mesotrophic", "Eutrophic"), names_to = "Trophic_State")

trending_oligotrophic_example$Trophic_State <- factor(trending_oligotrophic_example$Trophic_State, levels = c('Eutrophic', "Mesotrophic", "Oligotrophic"))


trending_oligotrophic_example_plot <- ggplot(trending_oligotrophic_example, aes(x = year, y = value, fill = Trophic_State)) +
  geom_bar(stat = "identity", show.legend = FALSE, color = 'black', width = 1) +
  scale_fill_manual(values = c( '#7dae38',  '#698c86', '#2158bc')) +
  labs(title = "B) Increasing in % Oligotrophic (Tinmath Reservoir, CO)") +
  xlab("Year") +
  ylab("% Occurence") +
  theme_bw() +
  theme(legend.title = element_blank(), axis.title = element_text(size = 30), axis.text = element_text(size = 24, color = 'black'), legend.text = element_text(size = 30), title = element_text(size = 30, face = 'bold'))

trending_oligotrophic_example_plot

## increasing eutrophic

trending_eutrophic_example <- trending_eutrophic %>%
filter(Hylak_id == 1054694) # Myeres Reservoir Wyoming
1041710
trending_eutrophic_example <- yearly_observations %>%
  filter(Hylak_id %in% trending_eutrophic_example$Hylak_id) %>%
  rename(Oligotrophic = "pct_oligotrophic", 
         Mesotrophic = "pct_mesotrophic",
          Eutrophic = "pct_eutrophic") %>%
  pivot_longer(cols = c("Oligotrophic", "Mesotrophic", "Eutrophic"), names_to = "Trophic_State")

trending_eutrophic_example$Trophic_State <- factor(trending_eutrophic_example$Trophic_State, levels = c('Eutrophic', "Mesotrophic", "Oligotrophic"))


trending_eutrophic_example_plot <- ggplot(trending_eutrophic_example, aes(x = year, y = value, fill = Trophic_State)) +
  geom_bar(stat = "identity", color = 'black', width = 1) +
  scale_fill_manual(values = c( '#7dae38',  '#698c86', '#2158bc')) +
  labs(title = "C) Increasing in % Eutrophic (Myers Reservoir, WY)") +
  xlab("Year") +
  ylab("% Occurence") +
  theme_bw() +
  theme(legend.title = element_blank(), axis.title = element_text(size = 30), axis.text = element_text(size = 24, color = 'black'), legend.text = element_text(size = 30), title = element_text(size = 30, face = 'bold'), legend.position = "bottom")

trending_eutrophic_example_plot

example_trends_plot <- ggarrange(no_trend_example_plot, trending_oligotrophic_example_plot, trending_eutrophic_example_plot, nrow = 3, ncol = 1)

example_trends_plot

```

# Examine the distribution of intensity of trophic state change across lakes becoming more eutrophic and those becoming more oligotrophic (Figure S1)
```{r}

oligotrophic_intensity <- ggplot(trending_oligotrophic, aes(x=oligotrophic_delta)) + 
  geom_histogram(binwidth=20, fill = 'grey60', colour = 'black') + 
  labs(x = "Change in % Oligotrophic", y  = 'Frequency', title = 'A) Lakes Trending Oligotrophic') +
 theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.title = element_blank(), axis.title = element_text(size = 30), axis.text = element_text(size = 24, color = 'black'), legend.text = element_text(size = 30), title = element_text(size = 30), legend.position = "bottom")

eutrophic_intensity <- ggplot(trending_eutrophic, aes(x=eutrophic_delta)) + 
  geom_histogram(binwidth=20, fill = 'grey60', colour = 'black') + 
  labs(x = "Change in % Eutrophic", y  = 'Frequency', title = 'A) Lakes Trending Oligotrophic') +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.title = element_blank(), axis.title = element_text(size = 30), axis.text = element_text(size = 24, color = 'black'), legend.text = element_text(size = 30), title = element_text(size = 30), legend.position = "bottom")

Figure_S1 <- ggarrange(oligotrophic_intensity, eutrophic_intensity, nrow = 2)

Figure_S1
```
